\Huge \bf \sc Calibration of an inertial measurement unit

### 2013-07-30 latex/paper_sensores/paper.tex
correcciones varias

Based on the assumption that both sensors have a linear response and considering the non-orthogonality of the axis of the devices we are going to present a standard model for both sensors(see \cite{bib:calib_imu}, \cite{bib:kalman}, \cite{bib:calib_imu_dos}).

@@

It is important to consider that the minimize function will generally have more than one local minimum. To ensure that the solution obtained is the desired one, the seed for the algorithm must be carefully chosen. Typically we will choose a vector $\boldsymbol{\theta}_0$ that is close to the values of the unknown parameters. Even if we do not know the exact value of the model's parameters, the datasheet of the different sensors can give us a very good clue about those values.\\

\subsubsection{Accelerometer}
\label{subsec:acc}
Since an accelerometer measures the acceleration in a free fall referential, in static equilibrium it will measure an acceleration that has as norm the gravitational constant of the  earth ($g = 9.81ms^{-2}$), its direction is collinear with the line defined by the position of the accelerometer and the center of the earth and its sense is from the center of the earth towards the position of the accelerometer, meaning ``up''. This allows us to have an exact expression for the acceleration, the procedure proposed consist in measure this magnitude with different orientations of the acceleration and then solve the problem stated in equation \refp{eq:calib_problem}.\\

@@

We are not going to use all the 27 measures to do the calibration, instead we use 14 of this measures as a training set and the other 13 as a test set. The 14 measures of acceleration give us 42 different acceleration values (3 for each measure corresponding to each axis), since we want to estimate 12 parameters we consider that is sufficient.\\

The delicate issue is knowing the exact orientation of the platform, having solve this is very easy to compute the theoretical acceleration that the device should measure in each axis from equations \refp{eq:acc_x}, \refp{eq:acc_y} and \refp{eq:acc_z}.\\

@@

As it was stated in \refp{sec:modelo} the problem \refp{eq:calib_problem} may have more than one local minimum, thus we need to assure that the seed of the algorithm is ``near'' of the values of the parameters. The proposed way to generate such a seed is to use some information of the sensor's datasheet\cite{bib:acc_data}:

@@

After this considerations the chose seed is:
\begin{equation}
{\theta}_0 = [26.10, 26.10, 26.10, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0]^T
\end{equation}
\subsubsection{Gyroscope}

@@

The mean error, and the standar deviation founded with the calibration method proposed in the test set are:

@@

\begin{equation}
\boldsymbol{\theta}_\omega = \begin{array}{c}
[792.44,\, 808.61, \, 799.44, \, -33.52,  \\
-31.41, \, -1.07, \, 0.02, \, 0.06, \\
 -1.32 \times 10^{-3}, \,  2.59\times 10^{2}, \,  -3.25 \times 10^{-2} \, 6.07\times 10^{-1}
 \end{array}]^T
\end{equation}

@@

The mean and the standard deviation obtained both correspond to aproximatelly $6$ bits errors.

As we can see the error obtained, in terms of bits, is bigger for the gyroscope than for the accelerometer. The main reason for this to happen is that in this case we are working with a scale that is not optimal, yet the only one available, for the speed of interest. The maximal angular speed measured ($198 ^{\circ} /s$ ) is approximately a tenth of the scale of the gyroscope ($2000^{\circ} /s$) while for the accelermoeter the maximal acceleration measured ($g$) is the half of the scale used ($2 g$).

@@

To evaluate the behavior of the measurement of the accelerometer and compensate the error caused by temperature, the experiment described in Section \refp{sec:param_temp} is carried out. It can be divided in two separate parts, the first is heating up the sensor using a hair dryer up to $48^oC$ and the second part is letting the sensor cool down to normal operating temperature. The two separated parts of the experiment are shown in Figure \refp{fig:temp_res}.

\begin{figure}[h!]
	\subfloat[Accelerometer measurement in\newline z-axis vs. temperature while heating\newline up the sensor.]{\label{fig:resultado_temp} 
			\includegraphics[width=.5\columnwidth]{./pics_paper/resultado_temp.png}}
	\subfloat[Accelerometer measurements and temperature (T) vs. time while T varying from $35^oC$ to $48^oC$.]{\label{fig:bajada}
		\includegraphics[width=.5\columnwidth]{./pics_paper/bajada.png}}
	\caption{\textbf{Temperature results -} Accelerometer measurements while temperature is modified.}
	\label{fig:temp_res}
\end{figure}

>>>>

Based on the assumption that both sensors have a linear response and considering the non-orthogonality of the axis of the devices we are going to present a standard model for both sensors (see \cite{bib:calib_imu}, \cite{bib:kalman}, \cite{bib:calib_imu_dos}).

@@

It is important to consider that the minimize function will generally have more than one local minimum. To ensure that the solution obtained is the desired one, the seed for the algorithm must be carefully chosen. Thus, datasheet values of the parameters are chosen as seed for the optimization.

Typically we will choose a vector $\boldsymbol{\theta}_0$ that is close to the values of the unknown parameters. Even if we do not know the exact value of the model's parameters, the datasheet of the different sensors can give us a very good clue about those values.\\

\subsubsection{Accelerometer}
\label{subsec:acc}
Since an accelerometer measures the acceleration in a free fall referential, in static equilibrium it will measure an acceleration that has as norm the gravitational constant of the  earth ($g = 9.81ms^{-2}$), its direction is collinear with the line defined by the position of the accelerometer and the center of the earth and its sense is from the center of the earth towards the position of the accelerometer, meaning ``up''. This allows to have an exact expression for the acceleration, the procedure used consist in measureing magnitude with different orientations of the acceleration and then solve the problem stated in equation \refp{eq:calib_problem}.\\

@@

The 27 measures are separated in two sets: 14 used for optimization and 13 as test set.\\

The key issue is knowing the orientation of the platform. Having solved this, it is easy to compute the theoretical acceleration that the device should measure in each axis from equations \refp{eq:acc_x}, \refp{eq:acc_y} and \refp{eq:acc_z}.\\

@@

As it was stated in \refp{sec:modelo} the problem \refp{eq:calib_problem} may have more than one local minimum, thus it is needed to assure that the seed of the algorithm is ``near'' of the values of the parameters. The proposed way to generate such a seed is to use some information of the sensor's datasheet\cite{bib:acc_data}:

@@

After this considerations the chosen seed is:
\begin{equation}
{\theta}_0 = [26.10, 26.10, 26.10, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0]^T
\end{equation}
\subsubsection{Gyroscope}

@@

The mean error, and the standard deviation founded with the calibration method proposed in the test set are:

@@

\begin{eqnarray}
\boldsymbol{\theta}_\omega &=& [792.44,\, 808.61, \, 799.44, \, -33.52, \, -31.41, \\ \nonumber
& & -1.07, \, 0.02, \, 0.06, \, -1.32 \times 10^{-3}, \, 2.59\times 10^{2}, \, \\ \nonumber
& & -3.25 \times 10^{-2}, \, 6.07\times 10^{-1}]^T \nonumber
\end{eqnarray}

@@

The mean and the standard deviation obtained both correspond to approximately $6$ bits errors.

As we can see the error obtained, in terms of bits, is bigger for the gyroscope than for the accelerometer. The main reason for this to happen is that in this case we are working with a scale that is not optimal, yet the only one available, for the speed of interest. The maximal angular speed measured ($198 ^{\circ} /s$ ) is approximately a tenth of the scale of the gyroscope ($2000^{\circ} /s$) while for the accelerometer the maximal acceleration measured ($g$) is the half of the scale used ($2 g$).

@@

To evaluate the behavior of the measurement of the accelerometer and compensate the error caused by temperature, the experiment described in Section \refp{sec:param_temp} is carried out. It can be divided in two separate parts, the first is heating up the sensor using a hair dryer up to $48^oC$ and the second part is letting the sensor cool down to normal operating temperature. The two separated parts of the experiment are shown in Figures \refp{fig:resultado_temp} and \refp{fig:bajada}.

\begin{figure}
	\centering
	\includegraphics[width=.8\columnwidth]{./pics_paper/resultado_temp.png}
	\caption{\textbf{Gyroscope setup - }Accelerometer measurement in z-axis vs. temperature while heating up the sensor.}
	\label{fig:resultado_temp}
\end{figure}
\begin{figure}
	\centering
	\includegraphics[width=.8\columnwidth]{./pics_paper/bajada.png}
	\caption{\textbf{Gyroscope setup - }Accelerometer measurements and temperature (T) vs. time while T varying from $35^oC$ to $48^oC$.}
	\label{fig:bajada}
\end{figure}
